Without loss of generality, \\ any reduced ring is a field

### 2018-05-23 paper-wlog.tex
Injectivity statement
Injective matrices

\title{Without loss of generality, \\ any reduced ring is a field}
\author{Ingo Blechschmidt}

@@

\newcommand{\XXX}[1]{\textbf{XXX: #1}}
\newcommand{\aaa}{\mathfrak{a}}
\newcommand{\bbb}{\mathfrak{b}}
\newcommand{\mmm}{\mathfrak{m}}
\newcommand{\ppp}{\mathfrak{p}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\defeq}{\vcentcolon=}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\Quot}{Quot}
\newcommand{\?}{\,{:}\,}
\newcommand{\notnot}{\emph{not not}\xspace}
\newcommand{\negg}{\neg\!\!\!\neg}

@@

>>>>

\title[Without loss of generality, any reduced ring is a field]{Without loss of generality, \\ any reduced ring is a field}
\author{Ingo Blechschmidt}

@@

\newcommand{\XXX}[1]{\textbf{XXX: #1}}
\newcommand{\aaa}{\mathfrak{a}}
\newcommand{\bbb}{\mathfrak{b}}
\newcommand{\mmm}{\mathfrak{m}}
\newcommand{\ppp}{\mathfrak{p}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\defeq}{\vcentcolon=}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\Quot}{Quot}
\newcommand{\?}{\,{:}\,}
\newcommand{\notnot}{\emph{not not}\xspace}
\newcommand{\negg}{\neg\!\!\!\neg}

@@

\begin{prop}\label{injection}
Let~$M \in A^{n \times m}$ be a matrix over a ring~$A$ such that~$n
< m$. Assume that the induced linear map~$A^m \to A^n$ is injective. Then~$1 =
0 \in A$.
\end{prop}

\begin{proof}[Proof~A (classical)]
Assume to the contrary that~$1 \neq 0 \in A$. Then there exists a minimal prime
ideal~$\ppp$. Since localization is exact, the induced linear map~$A_\ppp^m \to
A_\ppp^n$ is still injective. Since the stalk~$A_\ppp$ has precisely one prime
ideal, any element of~$A_\ppp$ is either invertible or nilpotent.

If~$A$ happens to be reduced, a shortcut is available: In this case~$A_\ppp$ is
also reduced, therefore a field. Thus injectivity of the linear map~$A_\ppp^m
\to A_\ppp^n$ is a contradiction to basic linear algebra.

In the general case, we argue as follows. Any entry of the matrix is invertible
or nilpotent as an element of~$A_\ppp$. If the former case occurs for at least
one entry, we can employ elementary row and column transformations to change it
into the unit, move it to the top left, kill the remaining entries in the first
row and in the first column, and continue in this fashion with the
resulting~$((n-1) \times (m-1))$-submatrix.

We eventually obtain a matrix of the
form~$\left(\begin{smallmatrix}I&0\\0&M'\end{smallmatrix}\right)$, where~$I$ is
an identity matrix of appropriate size, the submatrix~$M'$ is still injective,
and all entries of~$M'$ are nilpotent. Let~$\aaa$ be the ideal of~$A_\ppp$
generated by the entries of~$M'$. Since all generators of~$\aaa$ are nilpotent,
there is a natural number~$n \geq 0$ such that~$\aaa^n = (0)$.

If~$n \geq 1$, then in fact~$\aaa^{n-1} = (0)$, since if~$x \in \aaa^{n-1}$,
then~$M' \cdot (x,0,\ldots,0)^t = 0$ (the coefficients of the resulting vector are
elements of~$\aaa^n$) and thus~$x = 0$ by injectivity. Repeating this argument
yields~$(1) = \aaa^0 = (0)$, thus~$1 = 0 \in A_\ppp$; a contradiction.
\end{proof}

\begin{proof}[Proof B (constructive)]By XXX, the given matrix~$M$ is also
injective as a matrix over~$A^\sim$. We may therefore pass from~$A$ to the
forcing model~$A^\sim$, allowing us to assume that that~$A$ is almost a field
in the sense that any element is \notnot invertible or nilpotent.

Since the statement we want to verify,~$1 = 0 \? A^\sim$, is~$\neg\neg$-stable,
we may assume that any entry of the matrix is invertible or nilpotent. We can
therefore proceed exactly as in Proof~A, only referring to~$A^\sim$ where that
proof refers to the stalk~$A_\ppp$.
\end{proof}

\begin{prop}Let~$M \in A^{n \times m}$ be a matrix over a ring~$A$. Assume that
the induced linear map~$A^m \to A^n$ is injective. Then~$(\Lambda^m A)$, the
ideal of~$m$-minors of~$A$, is regular.
\end{prop}
